Compositional Verification of Concurrent C Programs with Search Structure Templates

### 2023-04-29 concurrency/CONCUR 2023/paper.tex
sketch proof of lock-coupling traverse

%maybe inline this instead of making it a figure?
\begin{figure}[ht]
	\centering
	\begin{mathpar}
		% spec of inRange
		{\color{blue}
			\forall \  m. \left\langle
			\nodeboxrep(\texttt{pn->n})  \ \ast \ \mathsf{R}(\texttt{pn->n}) \ \big| \ \treerep\ m \
			\right\rangle
		}
		\\ 
		\texttt{traverse(pn, k)} 
		\\
		{\color{blue}
			\left\langle \mathit{res}. \ \exists \  \texttt{n'}.
			\begin{array}{c}
				\ (\mathit{res} \leftrightarrow (\texttt{pn->n'->t} = \texttt{NULL}))  \ \ast \ \nodeboxrep(\texttt{pn->n'}) \\  
				\ast \ \mathsf{R}(\texttt{pn->n'})\ \ast \ (\texttt{k} \in \mathsf{range}(\texttt{n'}))
			\end{array}
			\ \Bigg| \ \treerep\ m \
			\right\rangle
		}
	\end{mathpar}
	\caption{Specification of \texttt{traverse} for the lock-coupling template}
	\label{fig:traverse_lock}
\end{figure}

@@

In the specification, the local precondition includes both the node handle $\nodeboxrep(\texttt{pn->n})$ and the contents $\mathsf{R}(\texttt{pn->n})$. The output of \lstinline{traverse} is a new node \lstinline{n'} in the data structure such that the key \texttt{k} falls within the range of \lstinline{n'}. The local postcondition then includes the handle of the new node $\nodeboxrep(\texttt{pn->n'})$ \wm{This looks weird: \lstinline{pn} doesn't have an \lstinline{n'} field}, its contents $\mathsf{R}(\texttt{pn->n'})$, and a Boolean variable $\mathit{res}$ indicating whether \texttt{pn->n'->t} is \texttt{NULL} or not \wm{Indicate the relationship between \lstinline{res} and the \lstinline{flag} return value}. The key to the correctness of the \texttt{traverse} function the loop invariant for the top-level loop, which expresses that in each iteration, \lstinline{traverse} holds the lock on a node that has \lstinline{k} in its range:
\begin{mathpar}\mathsf{traverse\_inv}(\texttt{pn}) \triangleq\ \exists \ \texttt{n'}.\ \texttt{k} \in \mathsf{range}(\texttt{n'})\ \land \ \nodeboxrep(\texttt{pn->n'})  \ \ast \  \mathsf{R}(\texttt{pn->n'})
\end{mathpar}\wm{Where does \lstinline{range} come from here?}

\newcommand\lstsetnumber[1]{\gdef\thelstnumber{#1}}
\begin{figure}[ht]
	$\color{blue}\left\langle\begin{array}{l} \nodeboxrep(b,g\_root)\ *\ \mathsf{atomic\_shift}(P_p,E_i,E_o,Q_p,Q)\end{array}\right\rangle$
	\begin{lstlisting}[language = C, style=myStyle, mathescape=true]
		int traverse(pn *pn, int k){
			  int flag = 1; 			$\left\{\begin{array}{l} 		\ghost{\mathsf{my\_half}((-\infty,+\infty),info)}{g\_root}\ *\ R\ b\ *\ \mathsf{lock\_inv}(l,lsh2,R')\ *\ \\
			  	\nodeboxrep(b,g\_root)\ *\ \mathsf{atomic\_shift}(P_p,E_i,E_o,Q_p,Q)\end{array}\right\} \Rrightarrow \left\{\begin{array}{l} lookup\_inv \end{array}\right\}$
		  		for( ; ; ){ 			$\left\{\begin{array}{l} lookup\_inv \end{array}\right\} \triangleq \left\{\begin{array}{l}(x\in range)\land \ghost{\mathsf{my\_half}(range,info)}{g\_current}*\ R\ np\ *\\\mathsf{lock\_inv}(lock,lsh2,R')\ *\ \nodeboxrep(b,g\_root)\ *\ \mathsf{atomic\_shift}(P_p,E_i,E_o,Q_p,Q)\end{array}\right\}$
		  			pn->p = pn->n;
		  			if (pn->p->t == NULL)
		  					break;
		  			else{
		  					int b = findNext(pn, k);
		  					if (b == 0){
		  							flag = 0;
		  							break;
		  					}
		  					else{
		  						acquire(pn->n->lock);
		  						release(pn->p->lock);
		  					}
		  				}
	  				}
  				return flag;
  		}  $\lstsetnumber{}$\end{lstlisting}
$\color{blue}\left\langle\begin{array}{l} \nodeboxrep(b,g\_root)\ *\ \mathsf{atomic\_shift}(P_p,E_i,E_o,Q_p,Q)\end{array}\right\rangle$
\caption{proof of \texttt{traverse}}
\end{figure}

>>>>

%maybe inline this instead of making it a figure?
\begin{figure}[ht]
	\centering
	\begin{mathpar}
		% spec of inRange
		{\color{blue}
			\forall \  m. \left\langle \texttt{pn} \mapsto (\texttt{p, n}) \ \ast \ 
			\nodeboxrep(\texttt{n})  \ \ast \ \mathsf{R}(\texttt{n}) \ \big| \ \treerep\ m \
			\right\rangle
		}
		\\ 
		\texttt{traverse(pn, k)} 
		\\
		{\color{blue}
			\left\langle \mathit{res}. \ \exists \  \texttt{n'}.
			\begin{array}{c} \texttt{pn} \mapsto (\texttt{p, n}) \ \ast 
				\ (\mathit{res} \leftrightarrow (\texttt{pn->n'->t} = \texttt{NULL}))  \ \ast \ \nodeboxrep(\texttt{pn->n'}) \\  
				\ast \ \mathsf{R}(\texttt{pn->n'})\ \ast \ (\texttt{k} \in \mathsf{range}(\texttt{n'}))
			\end{array}
			\ \Bigg| \ \treerep\ m \
			\right\rangle
		}
	\end{mathpar}
	\caption{Specification of \texttt{traverse} for the lock-coupling template}
	\label{fig:traverse_lock}
\end{figure}

@@

In the specification, the local precondition includes both the node handle $\nodeboxrep(\texttt{pn->n})$ and the contents $\mathsf{R}(\texttt{pn->n})$. The output of \lstinline{traverse} is a new node \lstinline{n'} in the data structure such that the key \texttt{k} falls within the range of \lstinline{n'}. The local postcondition then includes the handle of the new node $\nodeboxrep(\texttt{pn->n'})$ \wm{This looks weird: \lstinline{pn} doesn't have an \lstinline{n'} field}, its contents $\mathsf{R}(\texttt{pn->n'})$, and a Boolean variable $\mathit{res}$ indicating whether \texttt{pn->n'->t} is \texttt{NULL} or not \wm{Indicate the relationship between \lstinline{res} and the \lstinline{flag} return value}. The key to the correctness of the \texttt{traverse} function the loop invariant for the top-level loop, which expresses that in each iteration, \lstinline{traverse} holds the lock on a node that has \lstinline{k} in its range:
\begin{mathpar}\mathsf{traverse\_inv}(\texttt{pn}) \triangleq\ \exists \ \texttt{n'}.\ \texttt{k} \in \mathsf{range}(\texttt{n'})\ \land \ \texttt{pn} \mapsto (\texttt{p, n}) \ast \nodeboxrep(\texttt{n'})  \ \ast \  \mathsf{R}(\texttt{n'})
\end{mathpar}\wm{Where does \lstinline{range} come from here?}

\begin{figure}[h]
	$\color{blue}
	\forall \  m. \left\langle \texttt{pn} \mapsto (\texttt{p, n}) \ \ast \ 
	\nodeboxrep(\texttt{n})  \ \ast \ \mathsf{R}(\texttt{n}) \ \big| \ \treerep\ m \
	\right\rangle$
	\begin{lstlisting}[language = C, style=myStyle, mathescape=true]
		int traverse(pn *pn, int k){
			int flag = 1; 			$\left\{\begin{array}{l}  ...		\end{array}\right\} \Rrightarrow \left\{\begin{array}{l} \mathsf{traverse\_inv} \end{array}\right\}$
			for( ; ; ){ 			$\left\{\begin{array}{l} \mathsf{traverse\_inv} \end{array}\right\} \triangleq \left\{\begin{array}{l} ... \end{array}\right\}$
				pn->p = pn->n;
				if (pn->p->t == NULL)
				break;
				else{
					int b = findNext(pn, k);
					if (b == 0){
						flag = 0;
						break;
					}
					else{
						acquire(pn->n->lock);
						release(pn->p->lock);
					}
				}
			}
			return flag;
	}  \end{lstlisting}
	$\color{blue}
	\left\langle \mathit{res}. \ \exists \  \texttt{n'}.
	\begin{array}{c} \texttt{pn} \mapsto (\texttt{p, n'}) \ \ast 
		\ (\mathit{res} \leftrightarrow (\texttt{n'->t} = \texttt{NULL}))  \ \ast \ \nodeboxrep(\texttt{n'}) \ \ast \
		\\ \mathsf{R}(\texttt{n'})\ \ast \ (\texttt{k} \in \mathsf{range}(\texttt{n'}))
	\end{array}
	\ \Bigg| \ \treerep\ m \
	\right\rangle$
	\caption{The \texttt{traverse} function in lock-coupling template annotated with separation logic specification}
\end{figure}
